Stateless Model Checking with Data-Race Preemption Points

### 2015-11-16 pldi-16/paper.tex
implement pt2 of joshua comments

We have presented \quicksand, an implementation of Iterative Deepening, a new technique for automating the choice of PPs during stateless model checking.
% , and \quicksand, a tool which implements this technique tailored to the \landslide~model checker.
\quicksand~incorporates \landslide's data-race analysis to create new PPs tailored specifically to the program under test,
and manages multiple \landslide~instances to test many different PP subsets in a given CPU budget, even when the full state space of all PPs would be computationally intractable.

Our evaluation shows that \quicksand~achieves better bug-finding results than either single-pass data-race detection or single-state-space model checking (SSS-MC) alone.
%We dramatically reduce false-positive data race reports,
%%by verifying the absence of bugs in their associated state spaces
%and find bugs caused by nondeterministic data-races in alternate thread interleavings missed entirely by single-pass analysis.
Compared to SSS-MC,
%, we find bugs faster in smaller subset state spaces, %% not true, or at least, not by very much
%when the maximal state space is too large to explore
we find new bugs with data-race PPs that the maximal state space would not expose at all.
Moreover, for tests where all state spaces with data-race PPs can be completed, we provide an equally strong verification as if every single instruction had been used as a PP.

>>>>

We have presented Iterative Deepening and \quicksand, a new technique and tool for automating the choice of PPs during stateless model checking.
% , and \quicksand, a tool which implements this technique tailored to the \landslide~model checker.
\quicksand~incorporates data-race analysis to create new PPs tailored specifically to the program under test,
and manages multiple model checker instances to test many different PP subsets in a given CPU budget, even when the full state space of all PPs would be computationally intractable.

Our evaluation shows that \quicksand~achieves better bug-finding results than either single-pass data-race detection or single-state-space model checking (SSS-MC) alone.
%We dramatically reduce false-positive data race reports,
%%by verifying the absence of bugs in their associated state spaces
%and find bugs caused by nondeterministic data-races in alternate thread interleavings missed entirely by single-pass analysis.
Compared to SSS-MC,
%, we find bugs faster in smaller subset state spaces, %% not true, or at least, not by very much
%when the maximal state space is too large to explore
we find new bugs with data-race PPs that the maximal state space would not expose at all.
Moreover, when all data-race PPs can be fully tested within the CPU budget, we provide an equally strong verification as if every single instruction had been used as a PP.

### 2015-11-16 pldi-16/proof/di.tex
implement pt2 of joshua comments

@@

In this section we provide (hopefully) intuitive summaries of both our proof goals,
for readers who might be satisfied without necessarily double-checking the proofs' internal structure.

>>>>

Recent work \cite{sparc-ssm} has proposed hardware techniques for detecting many classes of stale heap pointer accesses, including the one shown in Figure \ref{fig:recycle-bug}.
This approach could be combined with stateless model checking in future work to identify such bugs immediately,
rather than requiring Iterative Deepening to explore new state spaces corresponding to the data race.
However, if the {\tt malloc} call were in thread 1 instead of thread 2, the bug would still be nondeterministic, requiring stateless model checking to expose.

@@

In this section we provide (hopefully) intuitive summaries of both our proof goals.
%for readers not interested in double-checking the proofs' internal structure.

### 2015-11-16 pldi-16/related.tex
some more polish; clean up stale todos; widen the table
clean up stale TODOs
one more case of anonymizing
implement pt2 of joshua comments

% TODO: ", and related data race analyses based on ????"

@@

\landslide~uses many established model-checking techniques, dating back
% of course
to Verisoft, the original C model checker \cite{verisoft}.
%, and Eraser, the original data race detector \cite{eraser}.
%Our model checker
%\landslide~\cite{landslide} itself implements many techniques from prior work (\sect{\ref{sec:landslide}}).
%itself implements DPOR \cite{dpor},
%state space estimation \cite{estimation},
%and data-race detection \cite{eraser}.
We compare related tools by their treatment of shared-memory thread communication.

{\bf Synchronization events only.} CHESS \cite{chess} and dBug \cite{dbug-ssv} instrument the thread library API, and can preempt programs only during calls to this API.
Hence they will miss any bugs that require interleaving threads at instruction granularity during a data race. CHESS provides a data-race analysis to report any such violations of its concurrency model to the user, but does not incorporate data-race candidates as PPs in future tests.

@@

\subsection{Data Race Detection}

@@

% TODO: Figure out why they claim "Happens before produces NO false positives, only benign races".
% It seems impossible.
% But if true, it means either (a) they can somehow identify FRM DRs on the 1st pass, not needing to replay,
% or (b) reuse of memory by malloc is somehow outside their concurrency model.
Closer to our work, replay analysis \cite{recordreplaydrs} also suppresses false positives by testing multiple thread interleavings.
%after finding data race candidates.
This work compares the immediately resulting program states for differences,
preferring to err on the side of false positives.
RaceFuzzer \cite{racefuzzer} avoids false positives by requiring an actual failure be exhibited,
although it uses random schedule fuzzing rather than stateless model checking.
Note while these techniques can also classify our malloc-recycle candidates as false positives (\sect{\ref{sec:recycle}}),
they require replaying the threads in a new interleaving.
Moreover, \cite{portend} argues that accurate classification may require many re-executions,
%according to many pre- and post-race sequences,
which is tantamount to adding a new state space in \quicksand.
Our proof in \sect{\ref{sec:recycle}} allows eliminating this special case with no additional replay beyond what DPOR already requires.

@@

% \subsection{Other Concurrency Testing Approaches} % TODO: Well?
%
% blah blah pldi'15 symbiosis DSP

>>>>

@@

We build upon many established model-checking techniques, dating back
% of course
to Verisoft, the original C model checker \cite{verisoft}.
%, and Eraser, the original data race detector \cite{eraser}.
%Our model checker
%\landslide~\cite{landslide} itself implements many techniques from prior work (\sect{\ref{sec:landslide}}).
%itself implements DPOR \cite{dpor},
%state space estimation \cite{estimation},
%and data-race detection \cite{eraser}.
We compare related tools by their treatment of shared-memory thread communication.

{\bf Synchronization events only.} CHESS \cite{chess} and dBug \cite{dbug-ssv} instrument the thread library API, and can preempt programs only during calls to this API.
Hence, they will miss any bugs that require interleaving threads at instruction granularity during a data race. CHESS provides a data-race analysis to report any such violations of its concurrency model to the user, but does not incorporate data-race candidates as PPs in future tests.

@@

\subsection{Data Race Detection}
\label{sec:related-dr}

@@

Closer to our work, replay analysis \cite{recordreplaydrs} also suppresses false positives by testing multiple thread interleavings.
%after finding data race candidates.
This work compares the immediately resulting program states for differences,
preferring to err on the side of false positives.
RaceFuzzer \cite{racefuzzer} avoids false positives by requiring an actual failure be exhibited,
although it uses random schedule fuzzing rather than stateless model checking.
Note while these techniques can also classify our malloc-recycle candidates as false positives (\sect{\ref{sec:recycle}}),
they require replaying the threads in a new interleaving.
Moreover, \cite{portend} argues that accurate classification may require many re-executions,
%according to many pre- and post-race sequences,
which is tantamount to adding a new state space in \quicksand.
Our proof in \sect{\ref{sec:recycle}} allows eliminating this special case with no additional replay beyond what DPOR already requires.

@@

% \subsection{Other Concurrency Testing Approaches}
%
% blah blah pldi'15 symbiosis DSP
